Formalizing Stack Safety as a Security Property\ifcameraready\else\vspace{-0.75cm}\fi

### 2023-05-12 submission5/core.tex
some clarification of new text and nits

\paragraph*{Machine Model}
We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
and stack pointer. 
%, and a standard ABI.
%\sna{Not in our formalism (hardcoded +4 in the definition of WBCF notwithstanding.)
%  This assumption is made in our example and testing.}
We make no particular assumptions about the provenance of the machine code; in particular,
we do not assume the use of any particular compiler.
The machine may possibly be enhanced with % security
enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
we assume that the behavior of these mechanisms is incorporated into the basic
step semantics of the machine, but we keep the enforcement state unchanged when constructing variant states.
Failstop behavior by enforcement mechanisms is modeled as stepping to the same state
(and thus silently diverging.)
%% \sna{I think that the payload concept is too nitty-gritty for this section.
%%   The chief use of the payload is that it allows every value to be interpreted
%%   as an address. In the tag world, it happens to coincide with a natural instance
%%   of our variant-compatibility relation.}

@@

We now consider how {\tt f} might misbehave and violate desirable
stack safety properties associated with {\tt main}. To put the violations in a
security framework, suppose that {\tt f} is actually an attacker seeking
to leak {\tt secret}. It might do so in a number of ways, shown as snippets of
assembly code in \cref{fig:f}.
%
Leakage is most obviously viewed as a violation of {\tt main}'s {\it confidentiality}.
In \cref{subfig:direct}, {\tt f} takes an offset from the stack
pointer, accesses {\tt secret}, and directly outputs it. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
can instead return that value so that {\tt main} stores it to {\tt out},
as in \cref{subfig:indirect}.
%
Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive}
with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
(\cref{subfig:integrity}).
Attacks of this kind do not violate {\tt main}'s confidentiality, but
rather its {\it integrity}.
In \cref{subfig:WBCF}, the attacker arranges to return to the
wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
violating the program's {\it well-bracketed control flow} (\(\wbcf\).)
%
In \cref{subfig:WBCF2}, a different attack violates \(\wbcf\), this time
by returning to the correct program counter but with the wrong stack pointer.%
\footnote{We pad the last two variants with {\tt nop}s just so that all the
snippets have the same length, which keeps the step numbering uniform in~\cref{fig:exec1}.}

@@

While it would be simple to define integrity as ``all sealed elements retain their
values after the call,'' this would be stricter than necessary. Suppose that
a callee overwrites some data of its caller, but the caller never accesses that data
(or only does so after re-initializing it.) This would be harmless, with the callee
essentially using the caller's memory as scratch space, but the caller never seeing any change.

@@

Consider the state \(\mach\) after step 5, whose context is \((V_2,\sigma)\), with the attacker from
\cref{subfig:direct}. We take a variant state over the set of elements that are
\(\sealed\) in \(V_2\) (see \cref{fig:variant2}.)
If we take a trace of execution from each state until it returns,
the traces may differ, in this case outputting 5 (the original value of {\tt secret}) and
4 (its value in the variant) respectively. This is a violation of
{\it internal confidentiality} (formalized in \cref{tab:props}, line 3a).

@@

We implement and test two micro-policies from Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
{\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} (\(LTC\)).
They share a common structure: each function activation is assigned a ``color'' \(n\)
representing its identity. Stack locations belonging to that activation are
tagged \(\tagStackDepth{n}\), and while the activation is running, the tag on the
program counter (PC tag) is \(\tagPCDepth{n}\). Stack locations not part of
any activation are tagged \(\tagNoDepth\).

@@

\paragraph*{Lazy Applicability}

Roessler and DeHon describe {\em Lazy Tagging and Clearing} as aiming to enforce
a fundamentally different property than {\em Depth Isolation}. They consider \(DI\)
a form of memory safety, while they claim that {\em dataflow integrity} enforces a
form of {\em dataflow integrity}. Because our properties are phrased in
terms of dataflow, they represent a plausible definition  against which \(LTC\) can be tested.

\sna{Discuss the distinction that they make between memory safety and
  dataflow integrity; we have framed stack safety as a dataflow integrity
  problem. They did not give a formal specification, and their implementation
  does not enforce ours. Spell out the violation. Priority 1.}

\paragraph*{Blessed Sequences}

@@

@@

Roessler and DeHon do not protect registers, since the compiler is in the
trusted computing base, but we support threat models that do not.
In particular, \(\clri\) requires callee-saved
registers to be saved and restored properly. We extend \(DI\) and \(LTC\)
so that callee-saved registers, too, are tagged with the color of the
function that is using them. In \(DI\) they are tagged as part of the entry
sequence, while in \(LTC\) they are tagged when a value is placed in them.

@@

We can model this as a trio of additional security-relevant operations: one which
declares a register to contain a valid pointer, one which transmits the provenance
of a pointer from one element to another, and one which clears the provenance
(for instance, when a pointer is modified in place in a way that makes it invalid.)

>>>>

\paragraph*{Machine Model}
We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
and stack pointer. 
%, and a standard ABI.
%\sna{Not in our formalism (hardcoded +4 in the definition of WBCF notwithstanding.)
%  This assumption is made in our example and testing.}
We make no particular assumptions about the provenance of the machine code; in particular,
we do not assume the use of any particular compiler.
The machine may possibly be enhanced with % security
enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
we assume that the behavior of these mechanisms is incorporated into the basic
step semantics of the machine, but we keep the enforcement state unchanged when constructing variant states.
Failstop behavior by enforcement mechanisms is modeled as stepping to the same state
(and thus silently diverging).
%% \sna{I think that the payload concept is too nitty-gritty for this section.
%%   The chief use of the payload is that it allows every value to be interpreted
%%   as an address. In the tag world, it happens to coincide with a natural instance
%%   of our variant-compatibility relation.}

@@

We now consider how {\tt f} might misbehave and violate desirable
stack safety properties associated with {\tt main}. To put the violations in a
security framework, suppose that {\tt f} is actually an attacker seeking
to leak {\tt secret}. It might do so in a number of ways, shown as snippets of
assembly code in \cref{fig:f}.
%
Leakage is most obviously viewed as a violation of {\tt main}'s {\it confidentiality}.
In \cref{subfig:direct}, {\tt f} takes an offset from the stack
pointer, accesses {\tt secret}, and directly outputs it. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
can instead return that value so that {\tt main} stores it to {\tt out},
as in \cref{subfig:indirect}.
%
Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive}
with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
(\cref{subfig:integrity}).
Attacks of this kind do not violate {\tt main}'s confidentiality, but
rather its {\it integrity}.
In \cref{subfig:WBCF}, the attacker arranges to return to the
wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
violating the program's {\it well-bracketed control flow} (\(\wbcf\)).
%
In \cref{subfig:WBCF2}, a different attack violates \(\wbcf\), this time
by returning to the correct program counter but with the wrong stack pointer.%
\footnote{We pad the last two variants with {\tt nop}s just so that all the
snippets have the same length, which keeps the step numbering uniform in~\cref{fig:exec1}.}

@@

While it would be simple to define integrity as ``all sealed elements retain their
values after the call,'' this would be stricter than necessary. Suppose that
a callee overwrites some data of its caller, but the caller never accesses that data
(or only does so after re-initializing it). This would be harmless, with the callee
essentially using the caller's memory as scratch space, but the caller never seeing any change.

@@

Consider the state \(\mach\) after step 5, whose context is \((V_2,\sigma)\), with the attacker from
\cref{subfig:direct}. We take a variant state over the set of elements that are
\(\sealed\) in \(V_2\) (see \cref{fig:variant2}).
If we take a trace of execution from each state until it returns,
the traces may differ, in this case outputting 5 (the original value of {\tt secret}) and
4 (its value in the variant) respectively. This is a violation of
{\it internal confidentiality} (formalized in \cref{tab:props}, line 3a).

@@

We implement and test two micro-policies inspired by
Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}:
{\em Depth Isolation} (\(DI\)) and {\em Lazy Tagging and Clearing} (\(LTC\)).
Their precise connection to Roessler and DeHon's work is discussed below.
They share a common structure: each function activation is assigned a ``color'' \(n\)
representing its identity. Stack locations belonging to that activation are
tagged \(\tagStackDepth{n}\), and while the activation is running, the tag on the
program counter (PC tag) is \(\tagPCDepth{n}\). Stack locations not part of
any activation are tagged \(\tagNoDepth\).

@@

@@

\paragraph*{Applicability to Roessler \& DeHon~\cite{DBLP:conf/sp/RoesslerD18}:}

For Roessler and DeHon (henceforward \emph{R\&D}), Lazy Tagging and Lazy Clearing are both optimizations that
can be applied to their Depth Isolation policy. Our version of \(LTC\)
corresponds to Depth Isolation with both optimizations applied. How closely
do our properties correspond to the specifications of their policies?

R\&D differentiate between memory safety policies (without lazy optimization)
and {\em data-flow integrity} policies (with lazy optimization). Our properties
are phrased in terms of data flow, and we apply them to both optimized and non-optimized
Depth Isolation.
R\&D do not attempt to define explicit formal properties, but they do list the
behaviors that they expect their data-flow integrity policies to prevent, namely:
reads from sealed objects
(our \(\clrc\)), writes to sealed objects
if they are later read (our \(\clri\)), and reads
from deallocated objects (our \(\clec\)).
They also note that Lazy Clearing prevents uninitialized reads,
which corresponds roughly to our \(\clei\).

R\&D note a flaw in Depth Isolation: because function activations
are identified by depth, a dangling pointer into a stack frame might be usable
when a new frame is allocated at the same depth. Our testing does not discover
this flaw, because we do not test address-taken objects, but it discovers a
related flaw under Lazy Tagging and Clearing that does not require
an object's address to be taken. If an activation reads a location
that was previously written by an earlier activation at the same depth, it will
violate callee confidentiality. If that location was in a caller's frame,
it also violates caller integrity and confidentiality.

R\&D propose addressing the dangling-pointer issue by
tracking both the depth of the current activation and the static identity
of the active function. This would not eliminate all instances of this issue, but it
would require the confidentiality-violating activation to be of the same
function that wrote the data in the first place, which is a significantly higher bar.
We propose instead tracking every activation uniquely, which should eliminate the
issue entirely, as it does in our tests.

@@

R\&D do not need to protect registers, since they include the compiler in their
trusted computing base, but we target threat models that do not.
In particular, \(\clri\) requires callee-saved
registers to be saved and restored properly. We extend \(DI\) and \(LTC\)
so that callee-saved registers, too, are tagged with the color of the
function that is using them. In \(DI\) they are tagged as part of the entry
sequence, while in \(LTC\) they are tagged when a value is placed in them.

@@

We can model this as a trio of additional security-relevant operations: one which
declares a register to contain a valid pointer, one which transmits the provenance
of a pointer from one element to another, and one which clears the provenance
(for instance, when a pointer is modified in place in a way that makes it invalid).
